---
title: "Enrollment Distribution"
format: html
echo: false
editor: source
---

Today, Duke has a medium-sized undergraduate body, with about 6.8 thousand students. The last century has seen a significant increase in enrollment, and here we will break down these students by class to examine any changes over time or relations between the numbers for each class. What does the size of each class say about student enrollment and more importantly, student retention?

Currently, Duke has one of the highest freshman retention rates in the country, with about 98% of students returning for their sophomore year. This can be compared with the national average of 69% and the North Carolina average of 66%. As of 2017, Duke's four-year graduation rate was around 95%, and six-year graduation rate has remained around 94%-96% for the last decade. Retention and graduation rates are important indicators of student achievement as they give insight into how many students are able to actually complete their degrees on time, rather than simply enroll at the school.

```{r}
#| message = FALSE
library(ggplot2)
library(tidyverse)
library(dplyr)
library(kableExtra)
```

```{r}
#| message = FALSE
enrollment <- read_csv("~/idhfinal/data/enrollment.csv")
```

```{r}
enrollment <- enrollment |>
  mutate(Total = TotalEnroll,
         Freshman = FreshmanEnroll,
         Sophomore = SophomoreEnroll,
         Junior = JuniorEnroll,
         Senior = SeniorEnroll) |>
  select(1, 7:11)

```

```{r}
enrollment_long <- enrollment |>
 pivot_longer(
    cols = -"YearFall",
    names_to = "Class",
    values_to = "Number"
  )
```


```{r}
#| label: fig-enrollmentclass
#| fig-cap: "The number of undergraduate students in each class from the years 1929 to 1998 (Source: Office of the University Registrar records, Statistics 1925-1999, Boxes 30, 31, 32)"
#| fig-height: 10
#| fig-width: 10
enrollment_long |>
  ggplot(
    aes(x = YearFall, y = Number, color = Class)
  ) +
  geom_point() +
  geom_line() +
  labs(x = "Year",
       y = "Number of Students",
       title = "Undergraduate Enrollment by Class") +
  scale_color_discrete(breaks = c("Freshman", "Sophomore", "Junior", "Senior", 
                               "Total"))
```
Taking a look at @fig-enrollmentclass, we can examine the first and last years for which we have data on, and see what kinds of changes have occurred. In 1929, the number of students decreased for each class from freshman to senior year. This indicates consistent drop-out rates each year. The trend of gradually decreasing numbers by class appears to have continued until the 1950s, duringin which the numbers seem to have converged and started to vary once more in the 70s. The 1970s-1990s saw many erratic changes, but by the end of the 1990s it appears there were more upperclassmen than underclassmen, presumably due to an increase in transfer students and gap year students. As noted above,It is worth noting that there are some significant gaps in the data, which make it hard to assess the validity of the trends during these few decades. However, we are still able to still delineate the general pattern.

```{r}
#| label: fig-classproportion
#| fig-cap: "The proportions of undergraduate students by class from the years 1929 to 1998 (Source: Office of the University Registrar records, Statistics 1925-1999, Boxes 30, 31, 32)"
enrollment_long |>
  filter(Class != "Total") |>
  ggplot(
    aes(x = YearFall, y = Number, fill = factor(Class, levels=c("Freshman","Sophomore", "Junior", "Senior")))
  ) +
  geom_bar(position = "fill", stat = "identity") +
  labs(x = "Year",
       y = "Proportion",
       title = "Enrollment Proportions by Class",
       fill = "Class") +
  scale_fill_discrete(breaks = c("Freshman", "Sophomore", "Junior", "Senior"))
```
@fig-classproportion helps us visualize the changes in class proportions. We can see the increase in retention rates in the gradual evening-out of the four classes by the late 1990s. On the far left, the size of the freshman class far exceeds those of upper classes, some years coming close to making up 50% of the undergraduate body, but the classes even out significantly as we move towards the right.

Note:

There were limitations in the collection of this data beyond just the missing years. The reports in the Duke Archives included enrollment by semester, so the numbers used are from the fall semester reports each year. Additionally, the total enrollment each year is not an exact sum of the numbers recorded for each class, as I discounted the students recorded as "Special" or "Advanced Freshman". However, these numbers were marginal and would not greatly effect the results.
